---
title: "EDA"
author: "Hanan"
date: "11/9/2018"
output: html_document
---

Explore and Summarize Data
========================================================

#Introduction:
In this project will use one of the exploratory data analysis techniques to explore the dataset of Red Wine Quality by using R. [The Wine Quality description file](https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityInfo.txt) which describes the variables and their meanings and how the data was collected.

> **First, Let's take a look at the dataset**

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(dplyr)
library(gridExtra)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
wines <- read.csv('wineQualityReds.csv')
```

```{r echo=FALSE, Glimpse}
glimpse(wines)
```

```{r echo=FALSE, Summary}
summary(wines)
```

> **Observations:** As we can see, the dataset consists of 12 variables and 1599 observations. Eleven of the variables are numerical except for the quality. The quality variable represents as an ordered factor. Its range from 3 to 8 with 6 being the median.


# Univariate Plots Section

> In this part, will show all variables with univariate analysis with plots.

# Quality

```{r echo=FALSE, Quality}
summary(wines$quality)
ggplot(data = wines, aes(x = quality))+
  geom_bar(fill="#CC79A7")
```

# Alcohol

```{r echo=FALSE, alcohol}
summary(wines$alcohol)
qplot(x= alcohol, data=wines, fill=I("#0072B2"), binwidth = .1)
```


#Sulphates

```{r echo=FALSE, sulphates}
summary(wines$sulphates)
qplot(x= sulphates, data=wines, fill=I("#F0E442"), binwidth = .1)

```

#pH

```{r echo=FALSE, pH}
summary(wines$pH)
qplot(x= pH, data=wines, fill=I("#999999"), binwidth = .1)

```

#Density

```{r echo=FALSE, density}
summary(wines$density)
qplot(x= density, data=wines, fill=I("purple"), binwidth = .001)

```



#Total sulfur dioxide

```{r echo=FALSE, total.sulfur.dioxide}
summary(wines$total.sulfur.dioxide)
qplot(x= total.sulfur.dioxide, data=wines, fill=I("#CC79A7"), binwidth = 4)
```



#Free sulfur dioxide

```{r echo=FALSE, free.sulfur.dioxide}
summary(wines$free.sulfur.dioxide)
qplot(x= free.sulfur.dioxide, data=wines, fill=I("#0072B2"), binwidth = 2)

```




#Chlorides

```{r echo=FALSE, Chlorides}
summary(wines$chlorides)
qplot(x= chlorides, data=wines, fill=I("#F0E442"), binwidth = .02)

```



#Residual sugar

```{r echo=FALSE,Residual.sugar}
summary(wines$residual.sugar)
qplot(x= residual.sugar, data=wines, fill=I("#999999"), binwidth = 1)

```




#Citric acid

```{r echo=FALSE, citric.acid}
summary(wines$citric.acid)
qplot(x= citric.acid, data=wines, fill=I("purple"), binwidth = .05)

```




#Volatile acidity

```{r echo=FALSE, volatile.acidity}
summary(wines$volatile.acidity)
qplot(x=volatile.acidity , data=wines, fill=I("#CC79A7"), binwidth = 0.1)

```

 
#Fixed Acidity

```{r echo=FALSE, fixed.acidity}
summary(wines$fixed.acidity)
qplot(x= fixed.acidity, data=wines, fill=I("#0072B2"), binwidth = 1)

```

 
 
 

```{r}

```


# Univariate Analysis


### What is the structure of your dataset?
There are 1599 observations. All observation are basised on 11 features which are fixed.acidity, volatile.acidity, citric.acid, residual.sugar, chlorides, free.sulfur.dioxide, total.sulfur.dioxide, density, pH,sulphates, and alcohol.

### What is/are the main feature(s) of interest in your dataset?
The main feature in the data set is the quality.

### What other features in the dataset do you think will help support your \

I think alcohol content, pH and total acidity level (volatile.acidity, fixed.acidity, citric.acid) will determine quality.

### Did you create any new variables from existing variables in the dataset?
No, I didn't create a new variable in the data set.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

There are no unusual distributions, no missing attribute values. This dataset is totally tidy and no need to change the form of the data.

# Bivariate Plots Section


```{r echo=FALSE, Bivariate_Plots}
ggplot(data= wines, aes(x = quality, y = citric.acid)) + 
  geom_point(color = "#CC79A7", fill="blue",alpha=1/4, position = 'jitter')+geom_smooth(method = "lm")+
  ggtitle(' Qaulity vs.Citic acid')


```

We can see that a higher citric acid will higher the quality.

```{r echo=FALSE }
ggplot(data= wines, aes(x = quality, y = residual.sugar)) + 
  geom_point(color = "#CC79A7", fill="blue",alpha=1/4, position = 'jitter')+geom_smooth(method = "lm")+
  ggtitle(' Quality vs.Residual sugar')
```


It looks the residual sugar has a low impact on the quality of red wines.

```{r echo=FALSE }
ggplot(data= wines, aes(x = quality, y = pH)) + 
  geom_point(color = "#CC79A7", fill="blue",alpha=1/4, position = 'jitter')+geom_smooth(method = "lm")+
  ggtitle(' Quality vs.pH')
```


It looks the pH has a low impact on the quality of red wines.


```{r echo=FALSE}
ggplot(data= wines, aes(x = quality, y = alcohol)) + 
  geom_point(color = "#CC79A7", fill="blue",alpha=1/4, position = 'jitter')+geom_smooth(method = "lm")+
  ggtitle(' Quality vs.alcohol')

```

It looks the alcohol content has a high impact on the quality of red wines.


```{r echo=FALSE}
ggplot(data= wines, aes(x = quality, y =  sulphates)) + 
  geom_point(color = "#CC79A7", fill="blue",alpha=1/4, position = 'jitter')+geom_smooth(method = "lm")+
  ggtitle(' Quality vs. sulphates')
```



It looks the sulphates have a high impact on the quality of red wines.


```{r echo=FALSE}
ggplot(data= wines, aes(x = quality, y = fixed.acidity)) + 
  geom_point(color = "#CC79A7", fill="blue",alpha=1/4, position = 'jitter')+geom_smooth(method = "lm")+
  ggtitle(' Quality vs. fixed.acidity')
```

We can see that quality increased when fixed acidity has been increased.


```{r echo=FALSE }
ggplot(wines, aes(x = residual.sugar, y = density))+
  geom_point(color = "#CC79A7", alpha = 1/3)+
  ggtitle(' Residual sugar vs.density')
```



We can see that residual sugar increased when density has been increased.


```{r echo=FALSE}
ggplot(wines, aes(x = alcohol, y = density))+
  geom_point(color = "#CC79A7", alpha = 1/3)+
  ggtitle(' Alcohol vs.density')
```


The density decrease with increase in the Alcohol content.




# Bivariate Analysis


### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

I have found that the correlation coefficient is positive for quality with citric. acid, fixed. acidity, sulphates, alcohol content. Also, it is positive between residual sugar and density.


### Did you observe any interesting relationships between the other features \
I have found that pH and residual sugar have a low impact on the quality of red wines.
It is interesting to see the relation between the density and the alcohol and sugar content.
### What was the strongest relationship you found?

 The strongest relationship that I found in this dataset was between the quality and the alcohol content.
# Multivariate Plots Section

# Correlation Matrix

```{r echo=FALSE, Multivariate_Plots}

mydata <- cor(wines)
```


```{r echo=FALSE, Matrix}

library('corrplot') 
corrplot(mydata, method = "ellipse")
```




```{r echo=FALSE}
ggplot(data=wines, aes(x=alcohol, y=pH, color=quality)) + 
  geom_point(alpha = 0.5, size = 2, position = 'jitter')+ 
  ggtitle('Alcohol vs pH with quality')
```


We can observe the quality of wines is impacted by pH. For the range of pH between 3.2 and 3.6 the quality was better.

```{r echo=FALSE}
ggplot(data=wines, aes(x=alcohol, y=sulphates, color=quality)) + 
  geom_point(alpha = 0.5, size = 2, position = 'jitter')+ 
  ggtitle('Alcohol vs sulphates with quality')
```

The high-quality wine contains a high quantity of sulphates and alcohol.

```{r echo=FALSE}
ggplot(data=wines, aes(x=alcohol, y=citric.acid, color=quality)) + 
  geom_point(alpha = 0.5, size = 2, position = 'jitter')+ 
  ggtitle('Alcohol vs citric acid with quality')
```

The high-quality wine contains a high quantity of citric acid and alcohol.


```{r echo=FALSE}
ggplot(data=wines, aes(x=alcohol, y=volatile.acidity, color=quality)) + 
  geom_point(alpha = 0.5, size = 2, position = 'jitter')+ 
  ggtitle('Alcohol vs volatile acidity with quality')
```

The high-quality wines contain high alcohol with low volatile acidity.



# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

-We have seen how alcohol and volatile acidity relates to quality.

-The high-quality wines contain high alcohol with low volatile acidity.

-The high-quality wine contains a high quantity of alcohol and citric acid.

-The higher amounts of alcohol with low volatile acidity content yield the best quality of wines.

-Also, the high quantity of sulphates and alcohol make the best wines.


# Final Plots and Summary


### Plot One
```{r echo=FALSE, Plot_One}
ggplot(data= wines, aes(x = quality, y = alcohol)) + 
  geom_point(color = "#CC79A7", fill="blue",alpha=1/4, position = 'jitter')+geom_smooth(method = "lm")+
  ggtitle(' Quality vs.alcohol')
```

### Description One
From the scatterplot, we can see that the quality is increasing with the increase in alcohol content. They have a positive and strong relati

### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(data=wines, aes(x = sulphates)) + 
  geom_histogram(binwidth = .05, fill="#cc79a7") +
  facet_wrap(~quality) +
  ggtitle('Relationship between sulphates and quality') 
```

### Description Two

As we can see from the histogram the quality of wines increased when the sulphates increased.


### Plot Three

```{r echo=FALSE, Plot_Three}
ggplot(wines, aes(x=pH, y= fixed.acidity)) + geom_point(color="#cc79a7") + geom_smooth(method = lm, se = FALSE)

```

### Description Three


When the amount of fixed acidity decrease the quantity of pH increased. The relationship between them is negative.


# Reflection

The red wine dataset contains 1599 information red wines. By using R, I had tried to get a sense of what factors might affect the quality of the wine to make it best. As we have found to the high-quality wines contain high alcohol with low volatile acidity. The high alcohol with low volatile acidity makes best wines. Also, the higher amounts of alcohol with low volatile acidity content yield the best quality of wines. The relationship between alcohol and citric acid with quality was positive and strong. The interesting to observed in this dataset was the relationship between alcohol and sugar with the density of wines.

